---
title: '326.212 Final Project: Part 2'
output: html_document
---

## 2019-14500 최민준

```{r}
library(RSQLite)
library(tidyverse)
con <- dbConnect(SQLite(), "project.sqlite")
```

## Part 2. Basic Questions (150 pts)

### Q1. Monthly traffic in three airports (60 pts)

#### Q1-1

```{r}
airports <- tbl(con, "airports")

filter(airports, IATA == "SNA" | IATA == "SJC" | IATA == "SMF")
```

- SNA: John Wayne Airport-Orange County Airport
- SJC: Norman Y. Mineta San Jose International Airport
- SMF: Sacramento International Airport



#### Q1-2

```{r}
flights <- tbl(con, "flights")

airportcounts <- flights %>%
  filter(Dest == "SNA" | Dest == "SJC" | Dest == "SMF") %>%
  group_by(Year, Month, Dest) %>%
  summarise(count = n()) %>%
  collect()

airportcounts
```



#### Q1-3

```{r}
library(lubridate)

airportcounts <- airportcounts %>%
  mutate(Date = lubridate::make_date(Year, Month)) %>%
  arrange(Date)

airportcounts
```

#### Q1-4

```{r}
ggplot(data = airportcounts, mapping = aes(x = Date, y = count, color = Dest)) +
  geom_line() +
  geom_point()

```



#### Q1-5

- For San Jose International Airport,
```{r}
filter(airportcounts, Dest == "SJC") %>%
  arrange(desc(count)) %>%
  head(10)
```

- For Sacramento International Airport,
```{r}
filter(airportcounts, Dest == "SMF") %>%
  arrange(desc(count)) %>%
  head(10)
```

- For Orange County Airport,
```{r}
filter(airportcounts, Dest == "SNA") %>%
  arrange(desc(count)) %>%
  head(10)
```




### Q2. Finding reliable airlines (60 pts)

#### Q2-1

```{r}
delays <- flights %>%
  filter(Origin == "ORD" & Dest == "MSP") %>%
  filter(Year == 2015) %>%
  filter(Op_Unique_Carrier == "UA" | Op_Unique_Carrier == "DL" | Op_Unique_Carrier == "AA" | Op_Unique_Carrier == "MQ") %>%
  select(1, 2, 3, 9, 15) %>%  # selecting only necessary columns
  group_by(Year, Month, Day_of_Month, Op_Unique_Carrier) %>%
  summarise(avg_ArrDelay = mean(Arr_Delay), count = n()) %>%
  collect()

delays

```

#### Q2-2

```{r}
airlines <- tbl(con, "airlines") %>%
  collect()

delays <- left_join(delays, airlines, by = c("Op_Unique_Carrier" = "Code"))
colnames(delays)[7] <- "Airline"

delays %>%
  ggplot() +
  geom_density(mapping = aes(x = avg_ArrDelay, color = Airline))
```

- It seems that American Air is the most reliable airline, because its density plot is most leftward (smallest average arrival-delay).
- The density plot of Envoy Air is certainly weighted to the right side compared to the other lines. This means that Envoy Air shows a bigger average arrival delay, and thus we can conclude that Envoy Air is the least reliable airline among these.
- In fact, Envoy Air is a regional airline in USA, while the others are major international carriers.




### Q3. All flights (30 pts)

#### Q3-1

```{r}
count_mondays_2001 <- flights %>%
  filter(Year == 2001 & Day_Of_Week == 1) %>%
  group_by(Year, Month, Day_of_Month) %>%
  summarise(count = n()) %>%
  collect() %>%
  mutate(Date = make_date(Year, Month, Day_of_Month))

count_mondays_2001

ggplot(data = count_mondays_2001) +
  geom_point(aes(x = Date, y = count)) +
  geom_line(aes(x = Date, y = count))
```

```{r}
count_mondays_2001 %>%
  filter(count < 17000)
```
Let's first explain the reasons of slightly decreased counts.
- January 1st: New Years Day
- May 28th: Memorial Day
- September 3rd: Labor Day
These are holidays in America, and this is the reason of the slight drops of counts of flights.

- The most interesting pattern is that there was a significant drop on number of flights from mid-September.
- The reason of this would be "9/11 terror attack" on the United States, which was on September 11th, 2001.
Directly after this terrorist attack, the federal government closed airports, canceling thousands of flights.
Even when the airports reopened, airlines experienced at least a 30 percent reduction in demand during the initial shock period. This might caused the airlines to reduce the number of flights.

- Also, December 24th is Christmas Eve, when people stay home. There is a rapid decrease in number of flights.
- December 31st, the day before New Years Day, also shows a relatively small number of flights.



#### Q3-2

```{r}
count_mondays_2011 <- flights %>%
  filter(Year == 2011 & Day_Of_Week == 1) %>%
  group_by(Year, Month, Day_of_Month) %>%
  summarise(count = n()) %>%
  collect() %>%
  mutate(Date = make_date(Year, Month, Day_of_Month))

count_mondays_2011

ggplot(data = count_mondays_2011) +
  geom_point(aes(x = Date, y = count)) +
  geom_line(aes(x = Date, y = count)) +
  coord_cartesian(ylim = c(12000, 19000)) +
  scale_y_continuous(breaks = seq(12000, 18000, by = 2000))

```

- There are slight outliers, but no big rapid changes compared to 2001.

These are holidays, when people usually stay home.
- May 30th: Memorial Day
- July 4th: Independence Day
- September 5th: Labor Day

- November 28th shows a noticeable increase, because it is the first Monday after Thanksgiving (Nov. 24). Many people visit their families during Thanksgiving and come back home, or to work.


```{r}
dbDisconnect(con)
```
